
library(rmarkdown)
library(plotly)



---
title: "Discretized_Age_Time_Structure"
author: "Laurette"
date: "29 October 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Discretized Age/Time Structure

This script summerises major functions using a discretised approach in simulating the susceptible and infected populations an  given age window and time.
The key inputs to the functions include:

1.  Susceptible Net Attritation rate - The probability of remaining in the susceptible population uninfeceted and not deceased  
2.  Infected Net Attritation rate - The probability of not dying due to  death (i.e excess mortality or natural death)
2.  Age window  - The time or number time steps the initial population is to be propagated into the future 
3.  Pop size  -  The population size 
4.  Del - time step 
5.  Birth counts -total number of births 

The Net Attrition rate is the probability of surviving being infected and not dying at age $a$ and time $t$ i.e $\rho(a, t)$. The number of susceptibles aged $a$ at time $t$,  $S(a,t)$ is given by 

$$S(a,t) = \rho(a - 1,t - 1)\cdot S(a - 1, t - 1)$$
The Birth counts is given as a vetor of the total number of births aged 0 at time $t$, and $t$ is infact a vector of times i.e $S(0, t) = S(0, t : t + \delta\cdot n )$   and $n\in(0,1,2,...)$. 

```{r neccesary functions, include=FALSE}

Make_birth_counts <- function(t_1,  
                              t_2, 
                              delta, 
                              N_0){
  birthcounts = data.frame(Time  = seq(t_1, t_2, delta),
                          Susceptible_0_t = (seq(t_1, t_2, delta)/sum(seq(t_1, t_2, delta))) * N_0)
  
  return(birthcounts)
}

Make_birth_counts(t_1 = 1,  
                  t_2 = 5, 
                  delta = 1/12, 
                  N_0 = 10000)





Incidence_var_a <- function(t, conc = 0.05, agemin =15,  
                            agemax = 50,  agepeak= 25, 
                            Imin =0.01,  Ipeak =0.05, 
                            Ifin =0.02){
  #varying Inicdence 
  incidence = ifelse(t <= agemin, 0, 
                     ifelse(t <= agepeak, Imin + ((Ipeak - Imin)/(agepeak - agemin)) * (t - agemin),
                            ifelse(t <= agemax, Ipeak + ((Ifin - Ipeak )/(agemax - agepeak)) * (t - agepeak), 0)))
  
  return(incidence)
}



Excess_Mortality_var_a <- function(t, conc = 0.05, agemin =15,  
                                   agemax = 50,   
                                   exmin =0.01,  
                                   exfin =0.05){
  #varying mortality
  Ex_mort = ifelse(t <= agemin, 0, 
                   ifelse(t <= agemax, exmin + ((exfin - exmin)/(agemax - agemin)) * (t - agemin),
                          0))
  
  return(Ex_mort)
}



Backgrnd_Mortality_var <- function(t, age = 1 ,Mort_peak = 0.01){
  Mort =  Mort_peak
  return (Mort)
}


```
 Above are the required functions to implement the code below (calculation of birth counts, incidence, and mortality).
```{r Survival_Susceptibles}
Survival_Susceptibles <- function(Age,
                                   Time, 
                                   Incidence, 
                                   Mortality, 
                                   pop_size, 
                                   Del,
                                   Birth_counts = Make_birth_counts){

    Susceptible  = matrix(0, nrow = length(Time) + Age, ncol =  length(1:Age))
   
   for (aa in 1:Age){
   
     if (aa == 1) {
     
     intergrand <-  (function(x){Incidence(x, Time + x) + Mortality(x, Time + x)})
     prob_survival = diag(1 - (sapply(Time,function(t) integrate( intergrand, lower = Age - 1, upper = Age )$value)))
     
     
     Susceptible[(1:length(Time)) + 1, 1] =  prob_survival %*% matrix(Birth_counts(t_1 = min(Time), 
                                                                                               t_2 = max(Time), 
                                                                                               delta = Del, 
                                                                                               N_0 = pop_size)$Susceptible_0_t, 
                                                                                               nrow = length(Time))
   }else{
     
     intergrand <-  (function(x){Incidence(x, Time + x) + Mortality(x, Time + x)})
     
     prob_survival = diag(1 - (sapply(Time,function(t) integrate( intergrand, lower = aa - 1, upper = aa )$value)))
     
     
     Susceptible[(1:length(Time)) + aa, aa] =  prob_survival %*% matrix(Susceptible[(1:length(Time) + aa) - 1, aa - 1], byrow = T)
     }
  
   }
   return(Susceptible)
   
 }  

```
Based on this function is was deemed fit to not calculate the attrition rate within the simulation of the suceptibles but to provide a matrix of attrition rate.
Key differences is that the above function takes in incidence and mortality rates and calculates the probability of surviving in the sucseptibles uninfected and not dying.


Below we present the same version of the same function the:
1. Net_attrition_rate 
2. Susceptibles_pop

We first estimate the attrition rate and subsequently parse the net attrition rate as a variable in the susceptibles function.


```{r net attrition rate}
y = Net_Attritions_rate <- function(Age,
                                Time, 
                                Incidence, 
                                Mortality, 
                                Del
                                ){
  
  
  #function calculates the attrition rates i.e. prob of not getting 
  #infected and the probability of not dying for a specific age and time.  
  
  net_Attritions_rate  = matrix(0, nrow = length(Time) + Age, ncol =  length(1:Age))
  
  for (aa in 1:Age){
    
    if (aa == 1) {
      
      intergrand <-  (function(x){Incidence(x, Time + x) + Mortality(x, Time + x)})
      #prob_survival = 1 - (sapply(Time,function(t) integrate( intergrand, lower = Age - 1, upper = Age )$value))
      
      
      net_Attritions_rate[(1:length(Time)) + 1, 1] =  1 - (sapply(Time,function(t) integrate( intergrand, lower = Age - 1, upper = Age )$value))
      
    }else{
      
      intergrand <-  (function(x){Incidence(x, Time + x) + Mortality(x, Time + x)})
      
      #prob_survival = 1 - (sapply(Time,function(t) integrate( intergrand, lower = aa - 1, upper = aa )$value))
      
      
      net_Attritions_rate[(1:length(Time)) + aa, aa] =  1 - (sapply(Time,function(t) integrate( intergrand, lower = aa - 1, upper = aa )$value))
    }
    
  }
  
  net_Attritions_rate = cbind( c(rep(1, length(Time)), rep(0,( nrow(net_Attritions_rate) - length(Time)))), net_Attritions_rate)
 
  return(net_Attritions_rate)
  
}



```


Example on output from the call to net attrition rate function 

```{r call net attrition rate, echo=FALSE}
 Net_Attritions_rate(Age = 2,
                        Time = seq(1, 3, 1), 
                        Incidence = Incidence_var_a, 
                        Mortality = Backgrnd_Mortality_var
                        )



```

and below is the calculation of the susceptibles for a specific age window and time 

```{r Susceptible population}


Susceptibles_pop <- function(Age,
                             Time,
                             pop_size = 10000,
                             prob_survival = Net_Attritions_rate(Age = 5,
                                                                 Time = seq(1, 5, 1), 
                                                                 Incidence = Incidence_var_a, 
                                                                 Mortality = Backgrnd_Mortality_var),
                                  Del  = 1,
                                  Birth_counts = Make_birth_counts){
  
  Susceptible  = matrix(0, nrow = length(Time) + Age, ncol =  length(1:Age)+1)
  
  Susceptible[(1:length(Time)), 1] = matrix(Birth_counts(t_1 = min(Time), 
                                                         t_2 = max(Time), 
                                                         delta = Del, 
                                                         N_0 = pop_size)$Susceptible_0_t, 
                                                         nrow = length(Time))
  
  for (aa in 1:Age){
    
    if (aa == 1){
      #  the indexing is confusing
      
      Susceptible[(1:length(Time)) + aa, aa + 1] =  diag(prob_survival[(1:length(Time))+aa, aa + 1]) %*% Susceptible[(1:length(Time)), 1]
      
    }else{
      
      
      Susceptible[(1:length(Time)) + aa, aa +1] =  diag(prob_survival[(1:length(Time))+aa, aa+1]) %*% matrix(Susceptible[(1:length(Time) + aa) - 1, aa], byrow = T)
    }
    
  }
  return(Susceptible)
  
}  

```


Example on output from the call to Susceptibles population function 

```{r call Susceptible population, echo=FALSE}
 Susceptibles_pop(Age = 2,
                  Time = seq(1, 3, 1),
                  pop_size = 10000,
                  Net_Attritions_rate(Age = 2,
                                      Time = seq(1, 3, 1), 
                                      Incidence = Incidence_var_a, 
                                      Mortality = Backgrnd_Mortality_var),
                                       Del  = 1)



```


##Simulation (Infected Population)

The simulation of the infected at a given age $a$, time $t$ for time since infection $\tau= aa - aa_0$ is  in two parts 
1. simulation of the $I(a, t, \tau =0)$, and 
2. simulation of $I(a, t, \tau >0 )$






